The role of standing variation in geographic convergent adaptation

### 2015-04-15 standing-variation-paper.tex
added Ne bit in discussion

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage[authoryear]{natbib}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{hyperref}
%\usepackage{latexml} % for \iflatexml
\usepackage{color}
\usepackage[utf8]{inputenc}  % for special characters in authors' names

\newcommand{\gc}[1]{{\it\color{blue}#1}}
\newcommand{\mfp}[1]{{\it\color{red}#1}}

\newcommand{\plr}[1]{{\it\color{green}#1}}
\usepackage{lineno}  %%%% FOR AMNAT
\newcommand{\var}{\mathop{\mbox{Var}}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}

@@

\paragraph{How then do substitutions occur?}
If it is rare for gene flow to rapidly spread selected alleles across a species range, 
how then do selected alleles ever become fixed within species? 
Drift alone will act only slowly to sort variants within species into
divergence among species. 
\gc{Slight selective differences in the pleiotropic effects (or linked
  background) among convergent alleles could allow one allele to
  press into areas occupied by other alleles.}
Furthermore, repeated bouts of adaptation in particular genomic regions may act 
to push a subset of previously selected alleles to fixation across the
species range, through the spread of the genetic backgrounds on which
they arise.  
However, it seems likely that this is a slow process compared to the
initial rapid spread of selected alleles.
\
\paragraph{Speciation and extinction as phases of molecular evolution?}
One potential resolution is that many selected alleles achieve
fixation, not through their own species-wide spread, but rather
through subsequent large-scale changes in geographic range size
induced by extirpation of the species over parts of its range
\citep[see ][and references therein for how such a model could be constructed]{barton2013modelling}. 
Such drops in range size may fix, or radically change the species-wide frequency, 
of alleles previously restricted to small portion of a species range. 
Furthermore, many modes of speciation are proposed to occur through a
geographically-limited subset of populations forming the basis of new species, 
e.g.\ the splitting off of part of the range through a vicariance event 
or dispersal of a subset of individuals. 
In this case, speciation will cause geographic assortment of
polymorphic ancestral variation, again acting to fix variants within
newly formed species that were previously polymorphic across ancestral species ranges.

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage[authoryear]{natbib}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{hyperref}
%\usepackage{latexml} % for \iflatexml
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}  % for special characters in authors' names

\newcommand{\gc}[1]{{\it\color{blue}#1}}
\newcommand{\mfp}[1]{{\it\color{red}#1}}
\newcommand{\plr}[1]{{\it\color{Fuchsia}#1}}

\usepackage{lineno}  %%%% FOR AMNAT
\newcommand{\var}{\mathop{\mbox{Var}}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}

@@

\paragraph{How then do substitutions occur?}
If it is rare for gene flow to rapidly spread selected alleles across a species range, 
how then do selected alleles ever become fixed within species? 
Drift alone will act only slowly to sort variants within species into
divergence among species. 
\plr{Of course, effective population sizes estimated from genetic diversity
    are nearly always small enough that fixation of neutral alleles by drift
    is expected to take less time than the divergence time between species.
    However, this observation is somewhat circular,
    and as reviewed by \citet{leffler2012revisiting},
    we do not understand why the implied ``effective population sizes'' differ so much
    from census population sizes;
    selective processes (as described here) are likely part of the explanation \citep{corbettdetig2015natural}.
}
\gc{Slight selective differences in the pleiotropic effects (or linked
  background) among convergent alleles could allow one allele to
  press into areas occupied by other alleles.}
Furthermore, repeated bouts of adaptation in particular genomic regions may act 
to push a subset of previously selected alleles to fixation across the
species range, through the spread of the genetic backgrounds on which
they arise.  
However, it seems likely that this is a slow process compared to the
initial rapid spread of selected alleles.
\
\paragraph{Speciation and extinction as phases of molecular evolution?}
One potential resolution is that many selected alleles achieve
fixation, not through their own species-wide spread, but rather
through subsequent large-scale changes in geographic range size
induced by extirpation of the species over parts of its range
\citep[see ][and references therein for how such a model could be constructed]{barton2013modelling}. 
Such drops in range size may fix, or radically change the species-wide frequency, 
of alleles previously restricted to small portion of a species range. 
Furthermore, many modes of speciation are proposed to occur through a
geographically-limited subset of populations forming the basis of new species, 
e.g.\ the splitting off of part of the range through a vicariance event 
or dispersal of a subset of individuals. 
In this case, speciation will cause geographic assortment of
polymorphic ancestral variation, again acting to fix variants within
newly formed species that were previously polymorphic across ancestral species ranges.
